paper figures

### 2025-01-31 paper_figures.Rmd
clean-up

```{r FigS3X1, fig.width=4 * fig_scale, fig.height=1.75 * fig_scale, include=FALSE, eval=FALSE}
dualPCA <- function(X, Y) {
  library(irlba)
  # use CCA-like extension of dual PCA as described here: https://xinmingtu.cn/blog/2022/CCA_dual_PCA/
  #sv <- irlba(as.matrix(X) %*% as.matrix(t(Y)), nv=n)
  sv <- svd(as.matrix(X) %*% as.matrix(t(Y)))
  x <- sv$u %*% diag(sqrt(sv$d))
  row.names(x) <- row.names(X)
  colnames(x) <- paste0('dPC',seq(1,dim(x)[2]))
  y <- sv$v %*% diag(sqrt(sv$d))
  row.names(y) <- row.names(Y)
  colnames(y) <- paste0('dPC',seq(1,dim(y)[2]))
  list(x=x, y=y, sdev=sqrt(sv$d/(dim(X)[2]-1)))
}

CH12.X <- CH12.data[CH12.take, mouse.cols] %>% 
  dplyr::mutate_if(is.numeric, scale)
mouse.X <- mouse.data[mouse.take, mouse.cols] %>% 
  dplyr::mutate_if(is.numeric, scale)

dpca <- dualPCA(CH12.X, mouse.X)
dpcadat <- cbind(rbind(dpca$x, dpca$y),
                 data.frame(group=c(CH12.data[row.names(dpca$x),'Genotype'],
                                    mouse.data$material)))

ggplot(dpcadat,aes(x=dPC1,y=dPC2, color=group)) +
  geom_point(size=1) +
  geom_polygon(data = dpcadat %>%
                 dplyr::group_by(group) %>%
                 dplyr::slice(chull(dPC1, dPC2)), alpha = 0.1, aes(fill=group), linewidth=.1) +
  my_theme() +
  scale_fill_manual(values=c(genotype_colors,tissue_colors)) +
  scale_color_manual(values=c(genotype_colors,tissue_colors)) +
  theme(aspect.ratio=dpca$sdev[2]**2/dpca$sdev[1]**2,
        legend.spacing=unit(0,'mm')) +
  labs(x=paste0('dPC 1 (',100*signif(dpca$sdev[1]**2/sum(dpca$sdev**2),3),'%)'),
       y=paste0('dPC 2 (',100*signif(dpca$sdev[2]**2/sum(dpca$sdev**2),3),'%)'),
       color='',shape='',fill='')
```

```{r FigS3X2, fig.width=3 * fig_scale, fig.height=4 * fig_scale, include=FALSE, eval=FALSE}
tmp <- t(t(t(as.matrix(rbind(CH12.X, mouse.X))) %*% as.matrix(dpcadat[,c('dPC1','dPC2')])) / colSums(dpcadat[,c('dPC1','dPC2')]**2)) %>%
  as.data.frame() %>%
  tibble::rownames_to_column('col') %>%
  dplyr::left_join(columns, by='col') %>%
  dplyr::mutate(col=factor(col, levels=columns %>% 
                             dplyr::filter(col %in% mouse.cols) %>%
                             dplyr::pull(col)),
                col_name=gsub("_downsampled|_switch",'',col))

ggplot(tmp,aes(x=dPC1,y=dPC2, color=col, shape=col)) +
  geom_point(size=1) +
  #geom_text_repel(aes(label=col), size=1) +
  my_theme() +
  scale_color_manual(values=cols_shapes %>%
                       dplyr::pull(var_col,col),
                     labels=cols_shapes %>%
                       dplyr::pull(col_name)) +
  scale_shape_manual(values=cols_shapes %>%
                       dplyr::pull(var_ind,col),
                     labels=cols_shapes %>%
                       dplyr::pull(col_name)) +
  theme(legend.key.size=unit(1.5,'mm'),
        aspect.ratio=dpca$sdev[2]**2/dpca$sdev[1]**2,
        legend.spacing=unit(0,'mm'),
        legend.key.spacing.y=unit(0,'mm'),
        legend.position='bottom',
        legend.box='horizontal') +
  guides(color=guide_legend(ncol=2),
         shape=guide_legend(ncol=2)) +
  labs(x='dPC 1', y='dPC 2',
       color='',shape='',fill='',alpha='')
```

@@

```{r FigS3X5, fig.width=1. * fig_scale, fig.height=1.75 * fig_scale, include=FALSE}
batch.CV <- c()
for (batch in unique(CH12.meta$Batch)) {
  train <- CH12.meta$Batch != batch
  test <- CH12.meta$Batch == batch
  X <- CH12.data[, CH12.cols] %>%
    dplyr::mutate_if(is.numeric, function(x) { 
      s <- numeric(sum(train | test)); 
      s[train] <- (x[train]-mean(x[train]))/sd(x[train]);
      s[test] <- (x[test]-mean(x[test]))/sd(x[test]);
      s
      }) %>%
    as.matrix()

@@

```{r Fig4D, fig.width=2.25 * fig_scale, fig.height=2 * fig_scale} 
FB.data.pooled <- read.csv('data/FB_data_pooled.csv', header=1, row.names=1) %>%
  dplyr::mutate(num_variants_somatic = num_variants * (1 - frac_variants_germline),
                status=ifelse(Diagnosis %in% c('Control','Subclass','Atypical'), 'Control', 'CVID')) %>%
  tibble::column_to_rownames('Sample')
QP.data <- read.csv('data/QP_data.csv', header=1, row.names=1) %>%
  dplyr::mutate(Diagnosis=plyr::revalue(Diagnosis,c('control'='Control'))) %>%
  tibble::column_to_rownames('Sample')

FB_train_donors <- paste0('FB_',sprintf('%0.2d', 1:35))
human.meta <- rbind(FB.data.pooled %>%
                      dplyr::mutate(dataset=ifelse(grepl('FB',Donor),
                                                   ifelse(Sample_donor %in% FB_train_donors,'training','testing'),
                                                   'validation')) %>%
                      dplyr::select(Donor,Diagnosis,status,Timepoint, dataset,Sex,Age,complication, EUROclass, pct_B, pct_IgA, pct_IgG),
                    QP.data %>%
                      dplyr::mutate(dataset='validation',
                                    Timepoint='t1',
                                    complication=NA,
                                    EUROclass=NA,
                                    pct_B=NA,
                                    pct_IgA=NA,
                                    pct_IgG=NA,
                                    status=ifelse(Diagnosis=='Control','Control','CVID')) %>%
                      dplyr::select(Donor,Diagnosis,status,Timepoint,dataset,Sex,Age,complication, EUROclass, pct_B, pct_IgA, pct_IgG)) %>%
  dplyr::mutate(Diagnosis=factor(Diagnosis,levels=c('Control','Subclass','Atypical','CVID','ICOS','Kabuki','IgG4def',
                                                    'AID','TNFRSF13B',
                                                    'NFKB2','PI3K','ATM','BRCA1','LIG4','NIPBL')),
                status=factor(status, levels=c('Control','CVID')))

@@

another PCA with only subclass features

```{r FigS4C, fig.width=2.75 * fig_scale, fig.height=1.75 * fig_scale}
FB.cols.subclass <- grep('frac_clusters_S[MGA]+',colnames(FB.data.pooled),value=TRUE)
FB.X.subclass <- FB.data.pooled[FB.data.pooled$Sample_donor %in% FB_train_donors,FB.cols.subclass] %>%
  dplyr::mutate_if(is.numeric,scale)

pca.subclass <- prcomp(FB.X.subclass)
pcadat.subclass <- cbind(pca.subclass$x,human.meta[human.meta$dataset=='training',])

ggplot(pcadat.subclass,aes(x=PC1,y=PC2)) +
  geom_point(aes(color=Diagnosis),size=1) +
  geom_line(aes(group=Donor, color=Diagnosis)) +
  geom_polygon(data = pcadat.subclass %>%
                 dplyr::mutate(status=ifelse(Diagnosis=='Subclass', 'Subclass', as.character(status))) %>%
                 dplyr::group_by(status) %>%
                 dplyr::slice(chull(PC1, PC2)), alpha = 0.25, aes(fill=status)) +
  my_theme() +
  scale_fill_manual(values=diagnosis_colors, guide='none') +
  scale_color_manual(values=diagnosis_colors) +
  guides(color=guide_legend(override.aes = list(shape=15,size=2), ncol=1)) +
  theme(aspect.ratio=pca.subclass$sdev[2]**2/pca.subclass$sdev[1]**2,
        legend.position='right',
        legend.key.size=unit(1.5,'mm'),
        legend.key.spacing.y=unit(0,'mm'),
        legend.box='horizontal') +
  labs(x=paste0('PC 1 (',100*signif(pca.subclass$sdev[1]**2/sum(pca.subclass$sdev**2),3),'%)'),
       y=paste0('PC 2 (',100*signif(pca.subclass$sdev[2]**2/sum(pca.subclass$sdev**2),3),'%)'),
       color='',shape='',fill='')
```

(this PCA uses `r dim(FB.X.subclass)[2]` features)

@@

```{r FigS4X1, fig.width=.8 * fig_scale, fig.height=1 * fig_scale, include=FALSE}
set.seed(4)
sample.CV <- c()
FB.samples <- row.names(FB.data.pooled)[!(FB.data.pooled$Diagnosis %in% c('Subclass','Atypical'))]
for (k in seq(1,20)) {
  tr <- runif(length(FB.samples)) < .75
  te <- !tr
  X <- FB.data.pooled[FB.samples, selected.cols] %>%
    dplyr::mutate_if(is.numeric, function(x) {
      s <- numeric(sum(tr | te));
      s[tr] <- (x[tr]-mean(x[tr]))/sd(x[tr]);
      s[te] <- (x[te]-mean(x[te]))/sd(x[te]);
      s
      }) %>%
    #dplyr::mutate_if(is.numeric, scale) %>%
    as.matrix()

@@

FB.meta <- FB.data.pooled[FB.samples,] %>%
  dplyr::mutate(Sample_donor=gsub("FB_","FB",Sample_donor),
                dataset=ifelse(grepl('FB',Donor),
                               ifelse(Sample_donor %in% FB_train_donors,'training','testing'),
                               'validation')) %>%
  dplyr::select(Donor,Sample_donor,Diagnosis,dataset,Sex,Age,Timepoint,
                splenomegaly,lymphoproliferation,pct_B,pct_IgA,pct_IgG,complication)

@@

<!-- ```{r FigS4X2, fig.width=4 * fig_scale, fig.height=1.75 * fig_scale, include=FALSE, eval=FALSE} -->
<!-- all.cols <- intersect(colnames(CH12.X), colnames(human.X)) -->
<!-- dpca <- dualPCA(CH12.X[,all.cols], human.X[,all.cols]) -->

<!-- dpcadat <- cbind(rbind(dpca$x,dpca$y), -->
<!--                 data.frame(species=c(rep('mouse',dim(dpca$x)[1]), -->
<!--                                      rep('human',dim(dpca$y)[1])), -->
<!--                            Genotype=c(CH12.meta[,'Genotype'], -->
<!--                                       human.meta[,'diagnosis']))) -->

<!-- ggplot(dpcadat,aes(x=dPC1,y=dPC2, color=Genotype)) + -->
<!--   geom_point(aes(shape=species),size=.5) + -->
<!--   geom_polygon(data = dpcadat %>% -->
<!--                  dplyr::group_by(Genotype) %>% -->
<!--                  dplyr::slice(chull(dPC1, dPC2)), alpha = 0.1, aes(fill=Genotype), linewidth=.1) + -->
<!--   my_theme() + -->
<!--   scale_fill_manual(values=c(diagnosis_colors,genotype_colors)) + -->
<!--   scale_color_manual(values=c(diagnosis_colors,genotype_colors)) + -->
<!--   scale_shape_manual(values=c('human'=15,'mouse'=16)) + -->
<!--   guides(fill=guide_legend(ncol=2, override.aes = list(pch=c(rep(16,5),rep(15,14)))), -->
<!--          color=guide_legend(ncol=2, override.aes = list(pch=c(rep(16,5),rep(15,14)))), -->
<!--          shape='none') + -->
<!--   labs(x=paste0('dPC 1 (',100*signif(dpca$sdev[1]**2/sum(dpca$sdev**2),3),'%)'), -->
<!--        y=paste0('dPC 2 (',100*signif(dpca$sdev[2]**2/sum(dpca$sdev**2),3),'%)'), -->
<!--        color='',shape='',fill='') -->

<!-- ggplot(dpcadat,aes(x=dPC1,y=dPC2, color=Genotype)) + -->
<!--   geom_point(aes(shape=species),size=.5) + -->
<!--   geom_polygon(data = dpcadat %>% -->
<!--                  dplyr::group_by(Genotype) %>% -->
<!--                  dplyr::slice(chull(dPC1, dPC2)), alpha = 0.1, aes(fill=Genotype), linewidth=.1) + -->
<!--   my_theme() + -->
<!--   facet_wrap(~species, ncol=2) + -->
<!--   scale_fill_manual(values=c(diagnosis_colors,genotype_colors)) + -->
<!--   scale_color_manual(values=c(diagnosis_colors,genotype_colors)) + -->
<!--   scale_shape_manual(values=c('human'=15,'mouse'=16)) + -->
<!--   theme(legend.key.size=unit(1.5,'mm'), -->
<!--         aspect.ratio=dpca$sdev[2]**2/dpca$sdev[1]**2) + -->
<!--   guides(fill=guide_legend(ncol=2, override.aes = list(pch=c(rep(16,5),rep(15,14)))), -->
<!--          color=guide_legend(ncol=2, override.aes = list(pch=c(rep(16,5),rep(15,14)))), -->
<!--          shape='none') + -->
<!--   labs(x=paste0('dPC 1 (',100*signif(dpca$sdev[1]**2/sum(dpca$sdev**2),3),'%)'), -->
<!--        y=paste0('dPC 2 (',100*signif(dpca$sdev[2]**2/sum(dpca$sdev**2),3),'%)'), -->
<!--        color='',shape='',fill='') -->

<!-- dpcadat[,'group'] <- as.character(dpcadat$Genotype) -->
<!-- dpcadat[FB.meta$Sample_donor,'group'] <- as.character(class_df$class) -->
<!-- dpcadat[,'dataset'] <- 'CH12' -->
<!-- dpcadat[dpcadat$species=='human','dataset'] <- human.meta$dataset -->
<!-- ggplot(dpcadat,# %>% dplyr::filter(dataset!='validation'), -->
<!--        aes(x=dPC1,y=dPC2, color=group)) + -->
<!--   geom_point(aes(shape=species),size=.5) + -->
<!--   geom_polygon(data = dpcadat %>% -->
<!--                  #dplyr::filter(dataset!='validation') %>% -->
<!--                  dplyr::group_by(group) %>% -->
<!--                  dplyr::slice(chull(dPC1, dPC2)), alpha = 0.1, aes(fill=group), linewidth=.1) + -->
<!--   my_theme() + -->
<!--   facet_wrap(~species, ncol=2) + -->
<!--   scale_fill_manual(values=c(diagnosis_colors,genotype_colors,class_colors)) + -->
<!--   scale_color_manual(values=c(diagnosis_colors,genotype_colors,class_colors)) + -->
<!--   scale_shape_manual(values=c('human'=15,'mouse'=16)) + -->
<!--   theme(legend.key.size=unit(1.5,'mm'), -->
<!--         aspect.ratio=dpca$sdev[2]**2/dpca$sdev[1]**2) + -->
<!--   guides(fill=guide_legend(ncol=2), #, override.aes = list(pch=c(rep(16,5),rep(15,10)))), -->
<!--          color=guide_legend(ncol=2), #, override.aes = list(pch=c(rep(16,5),rep(15,10)))), -->
<!--          shape='none') + -->
<!--   labs(x=paste0('dPC 1 (',100*signif(dpca$sdev[1]**2/sum(dpca$sdev**2),3),'%)'), -->
<!--        y=paste0('dPC 2 (',100*signif(dpca$sdev[2]**2/sum(dpca$sdev**2),3),'%)'), -->
<!--        color='',shape='',fill='') -->
<!-- ``` -->

<!-- ```{r FigS4X3, fig.width=3 * fig_scale, fig.height=4 * fig_scale} -->
<!-- tmp <- t(t(t(as.matrix(rbind(CH12.X[,all.cols], human.X[,all.cols]))) %*% as.matrix(dpcadat[,c('dPC1','dPC2')])) / colSums(dpcadat[,c('dPC1','dPC2')]**2)) %>% -->
<!--   as.data.frame() %>% -->
<!--   tibble::rownames_to_column('col') %>% -->
<!--   dplyr::left_join(columns, by='col') %>% -->
<!--   dplyr::mutate(col=factor(col, levels=columns %>%  -->
<!--                              dplyr::filter(col %in% all.cols) %>% -->
<!--                              dplyr::pull(col)), -->
<!--                 col_name=gsub("_downsampled|_switch",'',col)) -->

<!-- cols_shapes <- tmp %>%  -->
<!--   dplyr::distinct(category, col, col_name) %>% -->
<!--   dplyr::mutate(var_col=cat_cols[as.character(category)]) %>% -->
<!--   dplyr::group_by(category) %>% -->
<!--   dplyr::mutate(var_ind=1:n()) %>% -->
<!--   dplyr::mutate(col=factor(col, levels=levels(tmp$col))) %>%  -->
<!--   dplyr::arrange(col) -->

<!-- ggplot(tmp,aes(x=dPC1,y=dPC2, color=col, shape=col)) + -->
<!--   geom_point(size=1) + -->
<!--   #geom_text_repel(aes(label=col), size=1) + -->
<!--   my_theme() + -->
<!--   scale_color_manual(values=cols_shapes %>% -->
<!--                        dplyr::pull(var_col,col), -->
<!--                      labels=cols_shapes %>% -->
<!--                        dplyr::pull(col_name)) + -->
<!--   scale_shape_manual(values=cols_shapes %>% -->
<!--                        dplyr::pull(var_ind,col), -->
<!--                      labels=cols_shapes %>% -->
<!--                        dplyr::pull(col_name)) + -->
<!--   theme(legend.key.size=unit(1.5,'mm'), -->
<!--         aspect.ratio=dpca$sdev[2]**2/dpca$sdev[1]**2, -->
<!--         legend.spacing=unit(0,'mm'), -->
<!--         legend.key.spacing.y=unit(0,'mm'), -->
<!--         legend.position='bottom', -->
<!--         legend.box='horizontal') + -->
<!--   guides(color=guide_legend(ncol=2), -->
<!--          shape=guide_legend(ncol=2)) + -->
<!--   labs(x='dPC 1', y='dPC 2', -->
<!--        color='',shape='',fill='',alpha='') -->
<!-- ``` -->

```{r FigS4X4, fig.width=6.75 * fig_scale, fig.height=3.5 * fig_scale}
features.here <- c('cluster_entropy_downsampled',
                   'top_clone_occupancy_downsampled',
                   'mean_cluster_size_downsampled',
                   'donor_complexity',
                   'donor_score_W',
                   'donor_score_WGCW',
                   'homology_fw',
                   'n_homology_switch',
                   'mean_length_SA',
                   'frac_breaks_intra',
                   'frac_breaks_inversions',
                   'frac_breaks_duplications')

tmp <- cbind(rbind(CH12.data[,features.here],
                   FB.data.pooled[,features.here],
                   QP.data[,features.here]),
             rbind(CH12.meta %>%
                     dplyr::mutate(Diagnosis=Genotype,
                                   dataset='CH12') %>%
                     dplyr::select(Diagnosis, dataset),
                   human.meta[,c('Diagnosis','dataset')])) %>%
  dplyr::filter(Diagnosis %in% c('WT','Trp53bp1 KO','Brca1 KO','Rif1 KO','Lig4 KO','Control','CVID','TNFRSF13B','ATM','BRCA1')) %>%
  tibble::rownames_to_column('sample') %>%
  gather(col,value,features.here) %>%
  dplyr::mutate(col=factor(gsub("_downsampled|_switch",'',col),
                           levels=gsub("_downsampled|_switch","",features.here)),
                group=droplevels(factor(Diagnosis, levels=c(names(genotype_colors), names(diagnosis_colors)))))

ggplot(tmp, aes(x=ifelse(dataset=='CH12',as.numeric(group),as.numeric(group) + 1),y=value,fill=group)) +
  geom_boxplot(outlier.shape=NA, lwd=.5) +
  geom_jitter(width=.1,height=0,size=.1) +
  stat_compare_means(comparisons=list(c(1,2),
                                      c(1,3),
                                      c(1,4),
                                      c(1,5)),
                     method='wilcox.test',label = "p.signif", hide.ns=TRUE, size=2, vjust=.5) +
  stat_compare_means(comparisons=list(c(7,8),
                                      c(7,9),
                                      c(7,10),
                                      c(7,11)),
                     method='wilcox.test',label = "p.signif", hide.ns=TRUE, size=2, vjust=.5) +
  facet_wrap(.~col,ncol=6,scales='free_y') +
  my_theme() +
  coord_cartesian(clip='off') +
  scale_fill_manual(values=c(diagnosis_colors,genotype_colors)) +
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=.5),
        legend.position='none') +
  scale_x_continuous(breaks=c(seq(1,5), seq(1,5) + 6),
                     labels=levels(tmp$group)) +
  labs(x='',y='', color='Diagnosis')
```

<!-- ```{r aggregated_FB_boxplots, fig.width=4.5 * fig_scale, fig.height=1.7 * fig_scale} -->
<!-- tmp <- FB.X.aggregated %>% -->
<!--   tibble::rownames_to_column('sample') %>% -->
<!--   dplyr::left_join(FB.meta %>% tibble::rownames_to_column('sample'), by='sample') %>% -->
<!--   dplyr::mutate(name=gsub('t','',gsub('FB_','',paste0(Donor,'_',Timepoint)))) %>% -->
<!--   dplyr::left_join(class_df,by='name') %>% -->
<!--   gather(col,value,c('diversity','isotypes','breaks','context','structural')) -->

<!-- ggplot(tmp, aes(x=class,y=value,fill=class)) + -->
<!--   geom_boxplot(outlier.shape=NA, lwd=.5) + -->
<!--   geom_jitter(width=.1,height=0,size=.5) + -->
<!--   stat_compare_means(comparisons=list(c('cluster1','cluster3'), -->
<!--                                       c('cluster2','cluster3'), -->
<!--                                       c('cluster1','cluster2')), -->
<!--                      method='wilcox.test',label = "p.signif", hide.ns=TRUE, size=2, vjust=.5) + -->
<!--   facet_wrap(.~col,ncol=length(features.here),scales='free_y') + -->
<!--   my_theme() + -->
<!--   coord_cartesian(clip='off') + -->
<!--   scale_fill_manual(values=c(diagnosis_colors,class_colors)) + -->
<!--   theme(axis.text.x=element_text(angle=45,hjust=1,vjust=1), -->
<!--         legend.position='none') + -->
<!--   labs(x='',y='category score', color='Diagnosis') -->
<!-- ``` -->

<!-- check correlation with measured B cell proportion in FACS -->

<!-- ```{r pct_B_cor, fig.width=3 * fig_scale, fig.height=3 * fig_scale} -->
<!-- FB.data.pooled %>% -->
<!--   gather(feature, value, selected.cols) %>% -->
<!--   dplyr::group_by(feature) %>% -->
<!--   dplyr::summarise(R=cor(value, pct_B, use='na.or.complete')) %>% -->
<!--   dplyr::left_join(columns, by=c('feature'='col')) %>% -->
<!--   dplyr::mutate(feature=gsub('_downsampled|_switch','',feature)) %>% -->
<!--   dplyr::arrange(R) %>% -->
<!--   dplyr::mutate(feature=factor(feature, levels=feature)) %>% -->
<!--   ggplot(aes(y=R, x=feature, fill=category)) +  -->
<!--   geom_col() + -->
<!--   my_theme() + -->
<!--   coord_flip() + -->
<!--   scale_fill_manual(values=cat_cols) + -->
<!--   labs(x='',y='correlation with % B cells') + -->
<!--   theme(legend.position=c(.2,.9)) -->

<!-- FB.data.pooled %>% -->
<!--   gather(feature, value, selected.cols) %>% -->
<!--   dplyr::group_by(feature) %>% -->
<!--   dplyr::summarise(R=cor(value, pct_IgA, use='na.or.complete')) %>% -->
<!--   dplyr::left_join(columns, by=c('feature'='col')) %>% -->
<!--   dplyr::mutate(feature=gsub('_downsampled|_switch','',feature)) %>% -->
<!--   dplyr::arrange(R) %>% -->
<!--   dplyr::mutate(feature=factor(feature, levels=feature)) %>% -->
<!--   ggplot(aes(y=R, x=feature, fill=category)) +  -->
<!--   geom_col() + -->
<!--   my_theme() + -->
<!--   coord_flip() + -->
<!--   scale_fill_manual(values=cat_cols) + -->
<!--   labs(x='',y='correlation with % IgA cells') + -->
<!--   theme(legend.position=c(.2,.9)) -->

<!-- FB.data.pooled %>% -->
<!--   gather(feature, value, selected.cols) %>% -->
<!--   dplyr::group_by(feature) %>% -->
<!--   dplyr::summarise(R=cor(value, pct_IgG, use='na.or.complete')) %>% -->
<!--   dplyr::left_join(columns, by=c('feature'='col')) %>% -->
<!--   dplyr::mutate(feature=gsub('_downsampled|_switch','',feature)) %>% -->
<!--   dplyr::arrange(R) %>% -->
<!--   dplyr::mutate(feature=factor(feature, levels=feature)) %>% -->
<!--   ggplot(aes(y=R, x=feature, fill=category)) +  -->
<!--   geom_col() + -->
<!--   my_theme() + -->
<!--   coord_flip() + -->
<!--   scale_fill_manual(values=cat_cols) + -->
<!--   labs(x='',y='correlation with % IgG cells') + -->
<!--   theme(legend.position=c(.2,.9)) -->

<!-- ``` -->

>>>>

@@

```{r FigS3X, fig.width=1. * fig_scale, fig.height=1.75 * fig_scale, include=FALSE}
batch.CV <- c()
for (batch in unique(CH12.meta$Batch)) {
  train <- CH12.meta$Batch != batch
  test <- CH12.meta$Batch == batch
  X <- CH12.data[, CH12.cols] %>%
    dplyr::mutate_if(is.numeric, function(x) { 
      s <- numeric(sum(train | test)); 
      s[train] <- (x[train]-mean(x[train]))/sd(x[train]);
      s[test] <- (x[test]-mean(x[test]))/sd(x[test]);
      s
      }) %>%
    as.matrix()

@@

```{r Fig4D, fig.width=2.25 * fig_scale, fig.height=2 * fig_scale} 
FB.data.pooled <- read.csv('data/FB_data_pooled.csv', header=1, row.names=1) %>%
  dplyr::mutate(num_variants_somatic = num_variants * (1 - frac_variants_germline),
                status=ifelse(Diagnosis %in% c('Control','Subclass','Atypical'), Diagnosis, 'CVID')) %>%
  tibble::column_to_rownames('Sample')
QP.data <- read.csv('data/QP_data.csv', header=1, row.names=1) %>%
  dplyr::mutate(Diagnosis=plyr::revalue(Diagnosis,c('control'='Control'))) %>%
  tibble::column_to_rownames('Sample')

FB_train_donors <- paste0('FB_',sprintf('%0.2d', 1:35))
human.meta <- rbind(FB.data.pooled %>%
                      dplyr::mutate(dataset=ifelse(grepl('FB',Donor),
                                                   ifelse(Sample_donor %in% FB_train_donors,'training','testing'),
                                                   'validation')) %>%
                      dplyr::select(Donor,Diagnosis,status,Timepoint, dataset,Sex,Age, pct_B, pct_IgA, pct_IgG),
                    QP.data %>%
                      dplyr::mutate(dataset='validation',
                                    Timepoint='t1',
                                    pct_B=NA,
                                    pct_IgA=NA,
                                    pct_IgG=NA,
                                    status=ifelse(Diagnosis=='Control','Control','CVID')) %>%
                      dplyr::select(Donor,Diagnosis,status,Timepoint,dataset,Sex,Age, pct_B, pct_IgA, pct_IgG)) %>%
  dplyr::mutate(Diagnosis=factor(Diagnosis,levels=c('Control','Subclass','Atypical','CVID','ICOS','Kabuki','IgG4def',
                                                    'AID','TNFRSF13B',
                                                    'NFKB2','PI3K','ATM','BRCA1','LIG4','NIPBL')),
                status=factor(status, levels=c('Control','Subclass','Atypical','CVID')))

@@

@@

```{r FigS4X, fig.width=.8 * fig_scale, fig.height=1 * fig_scale, include=FALSE}
set.seed(4)
sample.CV <- c()
FB.samples <- row.names(FB.data.pooled)[!(FB.data.pooled$Diagnosis %in% c('Subclass','Atypical'))]
for (k in seq(1,20)) {
  tr <- runif(length(FB.samples)) < .75
  te <- !tr
  X <- FB.data.pooled[FB.samples, selected.cols] %>%
    dplyr::mutate_if(is.numeric, function(x) {
      s <- numeric(sum(tr | te));
      s[tr] <- (x[tr]-mean(x[tr]))/sd(x[tr]);
      s[te] <- (x[te]-mean(x[te]))/sd(x[te]);
      s
      }) %>%
    #dplyr::mutate_if(is.numeric, scale) %>%
    as.matrix()

@@

FB.meta <- FB.data.pooled[FB.samples,] %>%
  dplyr::mutate(Sample_donor=gsub("FB_","FB",Sample_donor),
                dataset=ifelse(grepl('FB',Donor),
                               ifelse(Sample_donor %in% FB_train_donors,'training','testing'),
                               'validation')) %>%
  dplyr::select(Donor,Sample_donor,Diagnosis,dataset,Sex,Age,Timepoint,
                pct_B,pct_IgA,pct_IgG)

@@
